\documentclass[11pt]{article}

\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{geometry}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{float}
\usepackage{diagbox}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{tabularx}
\geometry{margin=1in}
\hypersetup{hidelinks}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}

\newcommand{\Zm}{\mathbb Z_m}
\newcommand{\Pzero}{P_0}
\newcommand{\bulk}{\mathrm{bulk}}   % generic (non-stall) step; see \S\ref{app:routeE-cycle} for definition
\newcommand{\keywordline}[1]{\par\noindent\textbf{Keywords}\par\noindent #1\par}

\title{Hamilton decompositions of the directed $3$-torus $D(m)$:\\
explicit constructions for all $m\ge 3$}
\author{Sanghyun Park\\ Yonsei University\\ Seoul, Republic of Korea}
\date{}

\begin{document}
\maketitle

\begin{abstract}
We study Hamilton decompositions of the directed three-dimensional torus
\[
D(m)=\mathrm{Cay}((\mathbb Z_m)^3,\{e_1,e_2,e_3\})\cong \vec C_m\square \vec C_m\square \vec C_m,
\]
where $\vec C_m$ is the directed $m$-cycle.
We prove that $D(m)$ admits a decomposition into three arc-disjoint directed Hamilton cycles for every $m\ge 3$.

For odd $m$, five Kempe swaps of the canonical arc-coloring produce the decomposition; explicit return-map formulas on a fixed plane show that each color class is Hamiltonian.
For even $m$, a sign-product invariant yields a parity obstruction ruling out all Kempe-from-canonical approaches.
We therefore introduce an explicit alternative direction assignment, Route~E, for all even $m\ge 6$, whose return maps are derived in closed form.
The boundary case $m=4$ is handled separately.
\end{abstract}
\keywordline{Cayley digraphs, directed Hamilton cycles, graph products, Hamilton decompositions, Kempe swaps}


\section{Introduction}

Fix an integer $m\ge 3$ and write $\mathbb Z_m=\mathbb Z/m\mathbb Z$.
Let $V=(\mathbb Z_m)^3$ and consider the directed graph $D(m)$ with arc set
\[
(i,j,k)\to(i+1,j,k),\qquad (i,j,k)\to(i,j+1,k),\qquad (i,j,k)\to(i,j,k+1),
\]
with all arithmetic modulo $m$.
Equivalently,
\[
D(m)=\mathrm{Cay}((\mathbb Z_m)^3,\{e_1,e_2,e_3\})\cong \vec C_m\square \vec C_m\square \vec C_m.
\]
A \emph{directed Hamilton decomposition} of $D(m)$ is a partition of its $3m^3$ arcs into three arc-disjoint directed Hamilton cycles.

Hamilton decompositions of Cartesian products of cycles have a long history.
For undirected products, Kotzig proved that $C_m\square C_n$ decomposes into Hamilton cycles~\cite{Kotzig1973}, and Foregger extended this to products of several cycles~\cite{Foregger1978}.
Broad surveys include Bermond~\cite{Bermond1978} and Alspach, Bermond, and Sotteau~\cite{AlspachBermondSotteau1990}.
For further results on Hamilton decompositions of graph products, see Stong~\cite{Stong1991}.

On the directed side, Trotter and Erd\H{o}s characterized when the product of two directed cycles is Hamiltonian~\cite{TrotterErdos1978}, and Curran and Witte extended this by showing that the Cartesian product of any three or more nontrivial (length $\ge 2$) directed cycles has a directed Hamilton cycle~\cite{CurranWitte1985}.
These results concern a single Hamilton cycle; the decomposition problem---partitioning all arcs into Hamilton cycles---was subsequently treated by Bogdanowicz~\cite{Bogdanowicz2017,Bogdanowicz2020} and Darijani, Miraftab, and Morris~\cite{DarijaniMiraftabMorris2022}.
As $D(m)$ is a Cayley digraph on an abelian group, our work also falls within the Cayley-graph Hamiltonicity tradition surveyed by Witte--Gallian~\cite{WitteGallian1984} and Curran--Gallian~\cite{CurranGallian1996}.
Two closely related results are especially relevant.
Bogdanowicz's decomposition results show that Cartesian products of directed cycles can be partitioned into cycles of a prescribed common length~\cite{Bogdanowicz2017}; our problem is the extremal case where each cycle must be Hamiltonian.
On the other hand, Darijani, Miraftab, and Morris show that for arc-disjoint Hamilton \emph{paths}, the three-factor product remains an open case in the current general theory~\cite{DarijaniMiraftabMorris2022}.
The symmetric three-dimensional torus $D(m)$ therefore sits at a natural boundary between existing Hamiltonicity results and full arc-decomposition.
The present paper resolves that decomposition problem.
The Kempe-swap recoloring operations central to our constructions are standard tools in edge-coloring and reconfiguration; see Mohar~\cite{Mohar2006} and Stiebitz--Scheide--Toft--Favrholdt~\cite{StiebitzEtAl2012}.

Our approach rests on three key ideas.
\begin{enumerate}[label=\textup{(\roman*)},leftmargin=2.5em]
\item A \emph{sign-product invariant} for Kempe swaps, a general combinatorial identity holding for any coloring in which each color class is a permutation.
From this invariant we derive a parity obstruction showing that the even case cannot be reached from the canonical arc-coloring by Kempe swaps alone.
\item A \emph{return-map reduction}: because every arc of $D(m)$ increases the layer function $S=i+j+k$ by $1$, Hamiltonicity on the full vertex set $V$ of size $m^3$ reduces to the cycle structure of a return map on the plane $P_0=\{S=0\}$ of size $m^2$.
\item A necessary \emph{case split}: the sign-product invariant forces the odd and even cases to use fundamentally different coloring strategies.
\end{enumerate}
For odd $m$, the return maps arising from the Kempe-swap construction are affine skew-products whose single-cycle property follows from elementary number theory.
For even $m$, the parity obstruction rules out all Kempe-from-canonical approaches, and the proof relies on an explicit direction assignment that breaks the canonical pattern on the layers $S\in\{0,1,2\}$, followed by a return-map analysis using three different first-return reductions.

Our main result is the following.

\begin{theorem}\label{thm:main-all}
For every integer $m\ge 3$, the arc set of $D(m)$ decomposes into three directed Hamilton cycles.
\end{theorem}

The proof splits into three cases: the odd case via an affine five-swap construction (Section~\ref{sec:affine}), the even case $m\ge 6$ via Route~E (Section~\ref{sec:even}; full return-map derivations in Appendices~\ref{app:routeE}--\ref{app:routeE-cycle}), and the boundary case $m=4$ via a finite witness (Appendix~\ref{app:m4}).

\begin{table}[t]
\centering
\footnotesize
\begin{tabularx}{\textwidth}{@{}l>{\raggedright\arraybackslash}X@{}}
\toprule
notation & meaning \\
\midrule
$V=(\mathbb Z_m)^3$ & vertex set of the directed torus $D(m)$ \\
$S(i,j,k)=i+j+k\pmod m$ & layer function; every legal arc increases $S$ by $1$ \\
$P_t=\{S=t\}$ & layer plane / return section \\
\bottomrule
\end{tabularx}
\caption{Core geometric notation for $D(m)$.}
\label{tab:notation}
\end{table}

\section{Colorings, Kempe swaps, and return maps}\label{sec:prelim}

\subsection{Arc-colorings and direction assignments}
We define a \emph{coloring} of $D(m)$ to be a triple of permutations $(f_0,f_1,f_2)$ on $V$ such that
\[
\{f_0(v),f_1(v),f_2(v)\}=\{v+e_1,\ v+e_2,\ v+e_3\}
\qquad\text{for every }v\in V.
\]
Equivalently, at each vertex this assigns a bijection from the color set $\{0,1,2\}$ to the three out-arcs $\{v\to v+e_1,\, v\to v+e_2,\, v\to v+e_3\}$.
The color-$c$ subgraph is the functional digraph of $f_c$ (i.e., the digraph with arc set $\{v\to f_c(v):v\in V\}$) and is therefore a disjoint union of directed cycles.
We say that color $c$ is \emph{Hamilton} if $f_c$ is a single $m^3$-cycle.

\begin{example}[Canonical coloring]\label{ex:canonical}
The \emph{canonical coloring} assigns the identity direction triple
$(d_0,d_1,d_2)=(0,1,2)$ at every vertex, so that
\[
f_0(i,j,k)=(i+1,j,k),\qquad f_1(i,j,k)=(i,j+1,k),\qquad f_2(i,j,k)=(i,j,k+1).
\]
Each color is then a disjoint union of $m^2$ directed $m$-cycles.
This coloring serves as the starting point for the odd-case construction (Section~\ref{sec:affine})
and as the baseline whose sign product determines the parity barrier (Corollary~\ref{cor:parity-barrier}).
\end{example}

We will also need a relaxed notion that drops the permutation requirement.

\begin{definition}[Direction assignment]\label{def:direction-assignment}
A \emph{direction assignment} on $D(m)$ is a triple $\delta=(d_0,d_1,d_2)$ with $d_c:V\to\{0,1,2\}$ such that for every vertex $v$ the triple $(d_0(v),d_1(v),d_2(v))$ is a permutation of $(0,1,2)$.
It induces maps
\[
g_c(v):=v+e_{d_c(v)}\qquad (c\in\{0,1,2\}).
\]
We call the direction assignment a \emph{coloring} precisely when each induced map $g_c$ is a permutation of $V$.
\end{definition}

As we will see in Section~\ref{sec:even}, Route~E will first be specified only as a direction assignment;
its verification as a coloring is proved later from the bijectivity of its return maps via Lemma~\ref{lem:return-validity}.

\subsection{Kempe maps and Kempe swaps}
Fix two colors $r\ne s$.
Given a coloring $(f_0,f_1,f_2)$, define the \emph{Kempe map}
\[
\tau_{r,s}:=f_s^{-1}\circ f_r.
\]
Its cycles are exactly the alternating $(r,s)$-colored cycles in the $2$-colored subdigraph on colors $r$ and $s$.

\begin{definition}[Kempe swap]
Let $X\subseteq V$ be a union of cycles of $\tau_{r,s}$.
The \emph{Kempe swap of colors $(r,s)$ on $X$} is the operation producing a new triple $(f'_0,f'_1,f'_2)$ by exchanging the color-$r$ and color-$s$ arcs at every vertex in $X$:
\[
f'_r(v)=
\begin{cases}
f_s(v), & v\in X,\\
f_r(v), & v\notin X,
\end{cases}
\qquad
f'_s(v)=
\begin{cases}
f_r(v), & v\in X,\\
f_s(v), & v\notin X,
\end{cases}
\qquad
f'_t=f_t\ (t\notin\{r,s\}).
\]
\end{definition}

\begin{lemma}[Kempe swaps preserve validity]\label{lem:kempe-valid}
If $(f_0,f_1,f_2)$ is a coloring and $X$ is a union of cycles of $\tau_{r,s}$, then the Kempe swap on $X$ produces another valid coloring.
\end{lemma}

\begin{proof}
On each $\tau_{r,s}$-cycle $C\subseteq X$, the color-$r$ and color-$s$ arcs restrict to two perfect matchings from the source set $C$ onto the common image set $f_r(C)=f_s(C)$.
Indeed, if $C=(v_0,v_1,\dots,v_{\ell-1})$, then $f_r(v_t)=f_s(v_{t+1})$ along the cycle.
Swapping the two colors on $C$ simply exchanges these matchings, so every vertex incident with $C$ still has indegree $1$ and outdegree $1$ in each of the two colors.
Outside $X$ nothing changes.
\end{proof}

\subsection{A parity barrier}
For a permutation $\pi$ on a finite set, let $\mathrm{sgn}(\pi)\in\{\pm1\}$ denote its sign.

\begin{theorem}[Sign-product invariant]\label{thm:sign-invariant}
Let $(f_0,f_1,f_2)$ be a coloring and let $(f'_0,f'_1,f'_2)$ be obtained by a Kempe swap between two colors.
Then
\[
\mathrm{sgn}(f'_0)\,\mathrm{sgn}(f'_1)\,\mathrm{sgn}(f'_2)
=
\mathrm{sgn}(f_0)\,\mathrm{sgn}(f_1)\,\mathrm{sgn}(f_2).
\]
\end{theorem}

\begin{proof}
It suffices to consider a swap between colors $r$ and $s$, since the third color is unchanged.
Because the swap support $X$ is a disjoint union of $\tau_{r,s}$-cycles and the sign is multiplicative, it suffices to treat a single cycle.
Let $C=(v_0,v_1,\dots,v_{\ell-1})$ be one cycle of $\tau_{r,s}=f_s^{-1}\circ f_r$ contained in $X$.
Define $\sigma\in S(V)$ to be the $\ell$-cycle $(v_0\,v_1\,\cdots\,v_{\ell-1})$ on $C$ and to fix $V\setminus C$ pointwise.
A direct check using $f_r(v_t)=f_s(v_{t+1})$ shows
\[
f'_r=f_r\circ\sigma^{-1},\qquad f'_s=f_s\circ\sigma.
\]
Therefore
\[
\mathrm{sgn}(f'_r)\mathrm{sgn}(f'_s)
=
\mathrm{sgn}(f_r)\mathrm{sgn}(\sigma^{-1})\,\mathrm{sgn}(f_s)\mathrm{sgn}(\sigma)
=
\mathrm{sgn}(f_r)\mathrm{sgn}(f_s).
\]
When $X$ is a union of several cycles, applying the argument cycle by cycle shows that the swap on each cycle preserves the sign product, so the global sign product is invariant.
\end{proof}

\begin{corollary}[Parity barrier for Kempe-from-canonical when $m$ is even]\label{cor:parity-barrier}
Let $m$ be even. Then no sequence of Kempe swaps starting from the canonical coloring can reach a Hamilton decomposition of $D(m)$.
\end{corollary}

\begin{proof}
A directed Hamilton cycle on $N=m^3$ vertices is an $N$-cycle permutation and therefore has sign $(-1)^{N-1}=-1$ when $m$ is even.
Hence any decomposition into three Hamilton cycles must have sign product $(-1)^3=-1$.
In the canonical coloring, each color is a disjoint union of $m^2$ directed $m$-cycles.
Each $m$-cycle has sign $(-1)^{m-1}=-1$ since $m$ is even, and the product of $m^2$ such signs is $(-1)^{m^2}=+1$ since $m^2$ is even.
Hence the total sign product is $(+1)^3=+1$.
The result now follows from Theorem~\ref{thm:sign-invariant}.
\end{proof}

\subsection{The layer function and return maps}
Define the layer function
\[
S(i,j,k):=i+j+k\pmod m.
\]
Every allowed arc increases $S$ by $1$, so for every coloring and every color $c$ one has $S(f_c(v))=S(v)+1$.

\begin{lemma}[Plane invariance of Kempe maps]\label{lem:plane-invariant}
For any coloring and any two colors $r\ne s$, the Kempe map $\tau_{r,s}=f_s^{-1}\circ f_r$ preserves $S$.
Consequently, every $\tau_{r,s}$-cycle is contained in a single plane $P_t:=\{S=t\}$.
\end{lemma}

\begin{proof}
Let $w=f_r(v)$. Then $S(w)=S(v)+1$.
If $u=\tau_{r,s}(v)=f_s^{-1}(w)$, then $f_s(u)=w$, so $S(w)=S(u)+1$ and therefore $S(u)=S(v)$.
\end{proof}

\begin{corollary}[Plane swaps are always Kempe-valid]\label{cor:plane-swaps}
For any $t\in\mathbb Z_m$, the plane $P_t$ is a union of cycles of $\tau_{r,s}$ for every pair $r\ne s$.
Hence swapping two colors on an entire plane is always a valid Kempe swap.
\end{corollary}

\begin{lemma}[Return maps control cycle structure]\label{lem:return}
Fix a coloring and a color $c$.
Let $P_0=\{S=0\}$ and let
\[
F_c:=f_c^m\big|_{P_0}:P_0\to P_0
\]
be the return map.
Then the directed cycles of $f_c$ are in bijection with the directed cycles of $F_c$, and the length of a cycle of $f_c$ is $m$ times the length of the corresponding cycle of $F_c$.
In particular, $f_c$ is Hamilton on $V$ if and only if $F_c$ is a single $m^2$-cycle on $P_0$.
\end{lemma}

\begin{proof}
Along an $f_c$-orbit, the value of $S$ increases by $1$ at every step, so the orbit meets $P_0$ exactly once every $m$ steps.
The induced motion on $P_0$ is therefore exactly the return map $F_c=f_c^m|_{P_0}$.
\end{proof}


\begin{lemma}[Validity via return maps]\label{lem:return-validity}
Let $f:V\to V$ be a directed color map such that every step increases the layer $S$ by $1$.
Write $P_t=\{S=t\}$ and let $g_t:P_t\to P_{t+1}$ be the induced layer maps.
If the return map
\[
R=g_{m-1}\circ\cdots\circ g_0:P_0\to P_0
\]
is bijective, then each $g_t$ is bijective.
Consequently $f$ is a permutation of $V$.
\end{lemma}

\begin{proof}
All layers have the same finite size $m^2$; this equal-size property is essential for the argument.
Since $R$ is bijective, it is injective.
If $g_0$ were not injective, then $R$ would not be injective.
Hence $g_0$ is injective and therefore bijective.

Now write
\[
g_{m-1}\circ\cdots\circ g_1 = R\circ g_0^{-1},
\]
so the shorter composition is again bijective.
Repeating the same argument inductively shows that every $g_t$ is bijective.
Therefore every vertex has indegree $1$ and outdegree $1$ in this color, so $f$ is a permutation of $V$.
\end{proof}


\section{An affine construction for odd $m$}\label{sec:affine}

\subsection{The affine supports}
Recall the canonical coloring from Example~\ref{ex:canonical}.

Let
\[
P_0:=\{S=0\},\qquad P_1:=\{S=1\},
\]
and define the affine lines
\[
L_0:=\{S=0,\ k=0\},\qquad L_1:=\{S=1,\ k=0\}.
\]

\begin{theorem}[Affine five-swap construction]\label{thm:5swap}
Start from the canonical coloring and perform the following five Kempe swaps:
\begin{enumerate}[leftmargin=2.2em]
\item swap colors $(0,1)$ on the line $L_0$;
\item swap colors $(0,2)$ on the plane $P_0$;
\item swap colors $(0,1)$ on the plane $P_0$;
\item swap colors $(0,1)$ on the line $L_1$;
\item swap colors $(0,2)$ on the plane $P_1$.
\end{enumerate}
This produces a valid coloring for every $m\ge 3$.
\end{theorem}

\begin{proof}[Proof of Theorem~\ref{thm:5swap}]
Steps (2), (3), and (5) are Kempe-valid by Corollary~\ref{cor:plane-swaps} because they are supported on entire planes.
For step (1), the coloring is canonical, and
\[
\tau_{0,1}=f_1^{-1}\circ f_0:(i,j,k)\mapsto(i+1,j-1,k)
\]
preserves both $S$ and $k$, acting on $L_0=\{S=0,k=0\}$ as a single $m$-cycle.
Hence (1) is Kempe-valid.

For step (4), note that steps (1)--(3) modify only tails in $P_0$.
If $v\in P_1$, then $f_0(v)$ lies in $P_2$.
Any color-$1$ arc landing in $P_2$ must originate in $P_1$, because every legal arc increases $S$ by $1$.
Hence $f_1^{-1}$ on the image of $P_1$ in $P_2$ depends only on the outgoing color-$1$ arcs from $P_1$, which remain canonical after steps (1)--(3).
Therefore $\tau_{0,1}$ acts on $P_1$ as $(i,j,k)\mapsto(i+1,j-1,k)$, so $L_1$ is a single $\tau_{0,1}$-cycle and step (4) is Kempe-valid.
\end{proof}

\begin{definition}[Bump maps]
For $v=(i,j,k)\in V$ define
\[
\mathrm{bump}_i(v)=(i+1,j,k),\qquad
\mathrm{bump}_j(v)=(i,j+1,k),\qquad
\mathrm{bump}_k(v)=(i,j,k+1),
\]
with all arithmetic modulo $m$.
\end{definition}

\begin{proposition}[Closed-form rule after the five swaps]\label{prop:odd-closed}
Let $m\ge 3$ and let $v=(i,j,k)\in V$ with $S=i+j+k\pmod m$.
After the five swaps of Theorem~\ref{thm:5swap}, the resulting maps $(f_0,f_1,f_2)$ are
\[
f_0(v)=
\begin{cases}
\mathrm{bump}_j(v) & \text{if } S=0 \text{ and } k\neq 0,\\
\mathrm{bump}_k(v) & \text{if } S=1,\\
\mathrm{bump}_i(v) & \text{otherwise,}
\end{cases}
\]
\[
f_1(v)=
\begin{cases}
\mathrm{bump}_k(v) & \text{if } S=0,\\
\mathrm{bump}_i(v) & \text{if } S=1 \text{ and } k=0,\\
\mathrm{bump}_j(v) & \text{otherwise,}
\end{cases}
\]
\[
f_2(v)=
\begin{cases}
\mathrm{bump}_j(v) & \text{if } S\in\{0,1\} \text{ and } k=0,\\
\mathrm{bump}_i(v) & \text{if } S\in\{0,1\} \text{ and } k\neq 0,\\
\mathrm{bump}_k(v) & \text{otherwise.}
\end{cases}
\]
\end{proposition}

\begin{proof}
Outside $P_0\cup P_1$ none of the five swaps acts, so the coloring remains canonical there.

On $P_0$, step (1) swaps colors $(0,1)$ on $L_0$ and does nothing on $P_0\setminus L_0$.
Thus after step (1) one has
\[
(f_0,f_1,f_2)=
\begin{cases}
(\mathrm{bump}_j,\mathrm{bump}_i,\mathrm{bump}_k), & \text{on }L_0,\\
(\mathrm{bump}_i,\mathrm{bump}_j,\mathrm{bump}_k), & \text{on }P_0\setminus L_0.
\end{cases}
\]
Step (2) swaps colors $(0,2)$ on all of $P_0$, so this becomes
\[
(f_0,f_1,f_2)=
\begin{cases}
(\mathrm{bump}_k,\mathrm{bump}_i,\mathrm{bump}_j), & \text{on }L_0,\\
(\mathrm{bump}_k,\mathrm{bump}_j,\mathrm{bump}_i), & \text{on }P_0\setminus L_0.
\end{cases}
\]
Step (3) swaps colors $(0,1)$ on all of $P_0$, yielding
\[
(f_0,f_1,f_2)=
\begin{cases}
(\mathrm{bump}_i,\mathrm{bump}_k,\mathrm{bump}_j), & \text{on }L_0,\\
(\mathrm{bump}_j,\mathrm{bump}_k,\mathrm{bump}_i), & \text{on }P_0\setminus L_0.
\end{cases}
\]
Since $L_0=\{S=0,k=0\}$, this is exactly the stated rule on $S=0$.

On $P_1$, the first three steps do nothing.
Step (4) swaps colors $(0,1)$ on $L_1$, so after step (4)
\[
(f_0,f_1,f_2)=
\begin{cases}
(\mathrm{bump}_j,\mathrm{bump}_i,\mathrm{bump}_k), & \text{on }L_1,\\
(\mathrm{bump}_i,\mathrm{bump}_j,\mathrm{bump}_k), & \text{on }P_1\setminus L_1.
\end{cases}
\]
Step (5) swaps colors $(0,2)$ on all of $P_1$, so finally
\[
(f_0,f_1,f_2)=
\begin{cases}
(\mathrm{bump}_k,\mathrm{bump}_i,\mathrm{bump}_j), & \text{on }L_1,\\
(\mathrm{bump}_k,\mathrm{bump}_j,\mathrm{bump}_i), & \text{on }P_1\setminus L_1.
\end{cases}
\]
Since $L_1=\{S=1,k=0\}$, this is exactly the stated rule on $S=1$.
\end{proof}

\begin{theorem}[Odd case]\label{thm:odd-hamilton}
If $m$ is odd, the coloring of Proposition~\ref{prop:odd-closed} decomposes the arc set of $D(m)$ into three directed Hamilton cycles.
\end{theorem}

\begin{proof}
By Lemma~\ref{lem:return} it suffices to show that each return map $F_c=f_c^m|_{P_0}$ is a single $m^2$-cycle on $P_0$.
Parameterize $P_0$ by $(i,k)\in(\mathbb Z_m)^2$ with $j\equiv -i-k$.
A direct step-count from Proposition~\ref{prop:odd-closed} gives
\[
F_0(i,k)=(i-2+\mathbf 1_{k=0},\,k+1),\qquad
F_1(i,k)=(i+\mathbf 1_{k=-1},\,k+1),
\]
\[
F_2(i,k)=(i+2-2\mathbf 1_{k=0},\,k-2),
\]
with all arithmetic modulo $m$.

For $F_0$, the second coordinate cycles by $k\mapsto k+1$ and the net $i$-increment over one full $k$-cycle is
\[
(-1)+(m-1)(-2)=-2m+1\equiv 1\pmod m.
\]
Hence $F_0^m$ fixes $k$ and increments $i$ by $1$, so $F_0$ is a single $m^2$-cycle.

For $F_1$, again $k\mapsto k+1$ is an $m$-cycle, and exactly once per $k$-cycle one has $k=-1$, so $F_1^m$ fixes $k$ and increments $i$ by $1$.
Hence $F_1$ is a single $m^2$-cycle.

For $F_2$, the second coordinate evolves by $k\mapsto k-2$.
Since $m$ is odd, $-2$ generates $\mathbb Z_m$, so $k$ cycles through all values and hits $k=0$ exactly once per $m$ steps.
Over one full $k$-cycle the net $i$-increment is
\[
2(m-1)\equiv -2\pmod m,
\]
so $F_2^m$ fixes $k$ and shifts $i$ by $-2$, which has order $m$ because $\gcd(2,m)=1$.
Hence $F_2$ is a single $m^2$-cycle.
Applying Lemma~\ref{lem:return} completes the proof.
\end{proof}

\begin{proposition}[Failure of color~$2$ for even $m$]\label{prop:even-shatter}
Let $m\ge 4$ be even and let $(f_0,f_1,f_2)$ be the affine coloring of Proposition~\ref{prop:odd-closed}.
Then $f_0$ and $f_1$ are Hamilton cycles, while $f_2$ decomposes into exactly $m+2$ directed cycles.
\end{proposition}

\begin{remark}
The failure is confined to a single color and is caused by the fact that $-2$ is not a unit modulo an even~$m$.
This diagnosis motivates the even-case strategy of modifying only the low-layer directions that feed into the return map for color~$2$.
\end{remark}

\begin{proof}[Proof of Proposition~\ref{prop:even-shatter}]
The formulas for $f_0$ and $f_1$ are unchanged from the odd case, so those two colors are still Hamilton.
For $f_2$, the return map is
\[
F_2(i,k)=(i+2-2\mathbf 1_{k=0},\,k-2).
\]
Because $m$ is even, the update $k\mapsto k-2$ preserves the parity of $k$, so $P_0$ splits into the odd-$k$ half and the even-$k$ half.

On the odd-$k$ half, the second coordinate runs through all odd residues with period $m/2$ and never hits $0$.
Hence over one full $k$-cycle the first coordinate gains $2\cdot(m/2)=m\equiv 0$, so every orbit there has length $m/2$.
Since the odd-$k$ half has size $m^2/2$, it contributes
\[
\frac{m^2/2}{m/2}=m
\]
cycles.

On the even-$k$ half, the second coordinate again has period $m/2$, but now it hits $0$ exactly once per such period.
Therefore over one full $k$-cycle the net change in $i$ is
\[
2\left(\frac m2-1\right)=m-2\equiv -2 \pmod m.
\]
So the induced return map after $m/2$ steps is translation by $-2$ on $\mathbb Z_m$, which has exactly $\gcd(m,2)=2$ cycles.
Hence the even-$k$ half contributes $2$ cycles.

Altogether $f_2$ decomposes into $m+2$ directed cycles.
\end{proof}


\section{A return-map design principle and Route~E}\label{sec:even}

We now turn to the even case.
The parity barrier of Corollary~\ref{cor:parity-barrier} rules out all Kempe-from-canonical approaches, so we adopt a direct construction.
The odd proof shows that Hamiltonicity can often be reduced to a skew-product return map on $P_0\cong (\mathbb Z_m)^2$; we begin by stating the underlying criterion.

\begin{lemma}[Skew-product return maps]\label{lem:skew-product}
Fix $m\ge 1$, a step $d\in \mathbb Z_m$ with $\gcd(d,m)=1$, and a function $\alpha:\mathbb Z_m\to\mathbb Z_m$.
Define
\[
F(i,k):=(i+\alpha(k),\,k+d).
\]
Then $F$ is a permutation and
\[
F^m(i,k)=(i+\Delta,\,k),\qquad \Delta:=\sum_{t\in\mathbb Z_m}\alpha(t).
\]
Consequently, every orbit of $F$ has length $m^2/\gcd(\Delta,m)$, so $F$ is a single $m^2$-cycle iff $\gcd(\Delta,m)=1$.
\end{lemma}

\begin{proof}
Since $d$ is a unit, $k\mapsto k+d$ is an $m$-cycle and $F$ is bijective with inverse $(i,k)\mapsto(i-\alpha(k-d),k-d)$.
Iterating gives
\[
F^t(i,k)=\Bigl(i+\sum_{r=0}^{t-1}\alpha(k+rd),\,k+td\Bigr).
\]
At $t=m$ the second coordinate returns to $k$ and the sum ranges over all residues of $\mathbb Z_m$, giving $\Delta$.
The orbit length follows from the order of translation by $\Delta$ on the first coordinate.
\end{proof}

For the affine odd construction, Lemma~\ref{lem:skew-product} explains exactly why color $2$ breaks when $m$ is even: the step in the second coordinate is $-2$, which is not a unit modulo an even $m$.
This isolates the even design problem: one must alter only the low layers in order to replace the nonunit drift in color~$2$ by a primitive two-dimensional bulk displacement, while keeping the set of exceptional vertices finite and geometrically explicit.

The sign-product barrier shows that, for even $m$, no Kempe-from-canonical derivation can possibly work.
Accordingly, we introduce the Route~E \emph{direction assignment} on the low layers $S\in\{0,1,2\}$.
All higher layers remain canonical.

For $v=(i,j,k)\in V$, write $S(v)=i+j+k\pmod m$.
A direction triple $(d_0,d_1,d_2)\in\{0,1,2\}^3$ means that color $c$ uses the arc $v\mapsto v+e_{d_c}$.
At this stage we specify only a direction assignment in the sense of Definition~\ref{def:direction-assignment}; the fact that the associated maps are permutations, and hence form a coloring, is proved later in Theorem~\ref{thm:routeE-even-main}.

\begin{definition}[Explicit Route~E direction assignment]\label{def:routeE}
Let $m\ge 6$ be even. Define a direction triple $\delta(v)=(d_0(v),d_1(v),d_2(v))$ as follows.
The layer-by-layer structure is summarized below; the formal rules appear in the subsequent paragraphs.

\smallskip
\begin{center}
\footnotesize
\begin{tabular}{@{}lll@{}}
\toprule
layer & rule & complexity \\
\midrule
$S\ge 3$ & canonical $(0,1,2)$ & none \\
$S=2$ & two branches, split on $j=0$ vs.\ $j\ne 0$ & $2$ triples \\
$S=1$ & two branches, split on $i=0$ vs.\ $i\ne 0$ & $2$ triples \\
$S=0$ & six branches, depending on $m\bmod 6$ & $\le 6$ triples \\
\bottomrule
\end{tabular}
\end{center}
\smallskip

\noindent
The design principle is as follows.
Layers $S\ge 3$ keep the canonical triple, so the return map on $P_0$ sees their contribution as a uniform translation.
Layers $1$ and $2$ introduce one branch each: they test whether a single coordinate vanishes after the preceding step, and swap two directions if so.
Layer $0$ carries the main splicing burden: its casework, which depends on $m\bmod 6$, arranges the finitely many defect points so that the resulting lane families are joined into a single cyclic order.

\paragraph{Layers $S\notin\{0,1,2\}$.}
Use the canonical triple $(d_0,d_1,d_2)=(0,1,2)$.

\paragraph{Layer $S=1$.}
Use
\[
(d_0,d_1,d_2)=
\begin{cases}
(1,0,2), & i=0,\\
(2,0,1), & i\ne 0.
\end{cases}
\]

\paragraph{Layer $S=2$.}
Use
\[
(d_0,d_1,d_2)=
\begin{cases}
(2,1,0), & j=0,\\
(0,1,2), & j\ne 0.
\end{cases}
\]

\paragraph{Layer $S=0$, case $m\equiv 0,2\pmod 6$.}
Use
\[
(d_0,d_1,d_2)=
\begin{cases}
(1,0,2), & v\in X_{102},\\
(0,2,1), & v\in X_{021},\\
(2,1,0), & v\in X_{210},\\
(0,1,2), & v=(m-2,1,1),\\
(2,0,1), & v=(m-2,2,0),\\
(1,2,0), & \text{otherwise.}
\end{cases}
\]
\smallskip\noindent
where the subscript on each set records the assigned direction triple:
\begin{align*}
X_{102}&:=\{(0,0,0)\}\cup\{(i,1,m-1-i):1\le i\le m-3\}\cup\{(m-1,2,m-1)\},\\
X_{021}&:=\{(0,1,m-1)\}\cup\{(i,m-i,0):1\le i\le m-3\}\cup\{(m-1,0,1)\},\\
X_{210}&:=\{(0,j,m-j):2\le j\le m-1\}\cup\{(1,0,m-1)\}.
\end{align*}

\paragraph{Layer $S=0$, case $m\equiv 4\pmod 6$.}
Use
\[
(d_0,d_1,d_2)=
\begin{cases}
(1,0,2), & v\in Y_{102},\\
(0,2,1), & v\in Y_{021},\\
(2,1,0), & v\in Y_{210},\\
(0,1,2), & v\in\{(1,1,m-2),(m-2,1,1)\},\\
(2,0,1), & v\in\{(1,m-1,0),(m-2,2,0)\},\\
(1,2,0), & \text{otherwise.}
\end{cases}
\]
where
\begin{align*}
Y_{102}&:=\{(0,0,0)\}\cup\{(i,1,m-1-i):2\le i\le m-3\}\cup\{(m-1,2,m-1)\},\\
Y_{021}&:=\{(0,1,m-1)\}\cup\{(i,m-i,0):2\le i\le m-3\}\cup\{(m-1,0,1)\},\\
Y_{210}&:=\{(0,j,m-j):2\le j\le m-1\}\cup\{(1,0,m-1)\}\\
&\hspace{2em}\cup\{(1,j,m-1-j):2\le j\le m-2\}\cup\{(2,0,m-2),(2,m-1,m-1)\}.
\end{align*}
This completes the definition: set $f_c(v)=v+e_{d_c(v)}$; these are the associated color maps whose $m$-step behaviour on $P_0$ is analyzed below.
\end{definition}

Once stated, the construction is verified purely combinatorially.
We call a step at which the return map departs from the uniform translation a \emph{stall}; the number of stalls per orbit is bounded independently of~$m$.

Below we develop the return-map framework (\S\ref{sec:framework}), prove color~$2$ in full (\S\ref{sec:even-r2}), and assemble the main even-case theorem.
Example~\ref{ex:m6-color2} illustrates the full machinery on the smallest admissible case $m=6$.

\subsection{Framework for the Route~E return maps}\label{sec:framework}

The appendices contain the detailed tables and the longer case-by-case bookkeeping, but the logical mechanism of the even proof can be stated in the main text.
The first input is a three-step transducer extracted from Definition~\ref{def:routeE}.

\begin{proposition}[Three-step transducer for the Route~E return maps]\label{prop:routeE-transducer}
Parameterize
\[
P_0=\{(i,j,k)\in \mathbb Z_m^3:i+j+k\equiv 0\}
\]
by
\[
v(i,k):=(i,-i-k,k),\qquad (i,k)\in \mathbb Z_m^2.
\]
For a color $c\in\{0,1,2\}$ and a start point $v(i,k)\in P_0$, write the first three color-$c$ directions on the layers $S=0,1,2$ as
\[
\omega_c(i,k)=(a_c,b_c,c_c)\in\{0,1,2\}^3.
\]
Then the return maps satisfy
\[
R_0(i,k)=\bigl(i+N_0(\omega_0)-3,\ k+N_2(\omega_0)\bigr),
\]
\[
R_1(i,k)=\bigl(i+N_0(\omega_1),\ k+N_2(\omega_1)\bigr),
\]
\[
R_2(i,k)=\bigl(i+N_0(\omega_2),\ k+N_2(\omega_2)-3\bigr),
\]
where $N_r(\omega)$ counts the occurrences of direction $r$ in the word $\omega$.
Moreover:
\begin{enumerate}[label=\textup{(\roman*)},leftmargin=2.5em]
\item for color $1$, one always has $b_1=0$ and $c_1=1$;
\item for colors $0$ and $2$, the layer-$1$ direction depends only on whether the post-step-$1$ $i$-coordinate equals $0$, and the layer-$2$ direction depends only on whether the post-step-$2$ $j$-coordinate equals $0$.
\end{enumerate}
Appendix~\ref{app:routeE-derivation} therefore reduces the derivation of the piecewise formulas for $R_0,R_1,R_2$ to a finite table evaluation on the layer-$0$ partition.
\end{proposition}

\begin{proof}
After the first three steps from $P_0$, the orbit lies on layer $S=3$.
All subsequent layers are canonical, so until the next return to $P_0$ the map follows direction $c$ for exactly $m-3$ further steps.
This gives the displayed displacement formulas.

Now write $j=-i-k$ on $P_0$.
After step $1$, the $i$-coordinate is
\[
i_1=i+\mathbf 1_{a_c=0}.
\]
By Definition~\ref{def:routeE}, on layer $S=1$ the color-$1$ direction is always $0$, the color-$0$ direction is $1$ if $i_1=0$ and $2$ otherwise, and the color-$2$ direction is $2$ if $i_1=0$ and $1$ otherwise.
So part~\textup{(i)} is immediate, and for colors $0$ and $2$ the second letter $b_c$ is determined by the single test $i_1=0$.

After step $2$, the $j$-coordinate is
\[
j_2=j+\mathbf 1_{a_c=1}+\mathbf 1_{b_c=1}.
\]
Again by Definition~\ref{def:routeE}, on layer $S=2$ the color-$0$ direction is $2$ if $j_2=0$ and $0$ otherwise, while the color-$2$ direction is $0$ if $j_2=0$ and $2$ otherwise.
Thus for colors $0$ and $2$ the third letter $c_c$ is determined by the single test $j_2=0$.
The appendix tables simply evaluate these tests on the finite partition of $P_0$ induced by the layer-$0$ rule.
\end{proof}

\begin{lemma}[Generic three-step words and bulk vectors]\label{lem:routeE-generic}
Consider the Route~E direction assignment.
Suppose a start point $(i,k)\in P_0\cong\Zm^2$ has the base layer-$0$ triple $(1,2,0)$ and that, along the first two steps of the orbit, neither test $i_1=0$ on layer $1$ nor $j_2=0$ on layer $2$ is triggered.
Then the three low-layer words are constant:
\[
\omega_0(i,k)=120,\qquad \omega_1(i,k)=201,\qquad \omega_2(i,k)=012.
\]
Consequently the return maps follow the bulk branches
\[
R_0(i,k)=(i-2,k+1),\qquad R_1(i,k)=(i+1,k+1),\qquad R_2(i,k)=(i+1,k-2).
\]
\end{lemma}

\begin{proof}
The base layer-$0$ triple $(1,2,0)$ means that on layer $0$ the colors $0,1,2$ use directions $1,2,0$, respectively.
If the post-step-$1$ test $i_1=0$ is not triggered, then the layer-$1$ triple is the generic one $(2,0,1)$.
If the post-step-$2$ test $j_2=0$ is not triggered, then the layer-$2$ triple is the generic one $(0,1,2)$.
Reading off the color-$c$ letters from these three generic triples gives the displayed words $120,201,012$.
The displacement formulas now follow immediately from Proposition~\ref{prop:routeE-transducer}.
\end{proof}

\begin{remark}[structure of the defect sets]\label{rem:routeE-design}
Lemma~\ref{lem:routeE-generic} explains the bulk of the construction: almost everywhere the low-layer word is constant, so the return maps are translations by the three displayed bulk vectors.
Non-generic behaviour can occur only for three reasons:
\begin{enumerate}[label=\textup{(\alph*)},leftmargin=2em]
\item the start point already lies on one of the special layer-$0$ defect lines or endpoint corrections;
\item after the first step, the test $i_1=0$ is triggered on layer $1$;
\item after the second step, the test $j_2=0$ is triggered on layer $2$.
\end{enumerate}
In the $(i,k)$ coordinates on $P_0$, each of these conditions is affine, which is why the defect support in Lemma~\ref{lem:routeE-finite-defect} consists of affine lines plus finitely many isolated boundary points.
The boundary-point exceptions are then tuned only to splice the long arithmetic \emph{lane families}---maximal runs of consecutive lanes that share the same bulk itinerary---produced by the bulk vectors into a single cyclic order.
The residue-class split $m\equiv 0,2\pmod{6}$ versus $m\equiv 4\pmod{6}$ appears only at this final splicing stage: for $m\equiv 0,2\pmod{6}$ the primary defect lines already connect the relevant lane families, whereas for $m\equiv 4\pmod{6}$ one additional affine splice family is needed for colors $1$ and $0$.
\end{remark}

The next lemma is the counting lemma used in all three color proofs.

\begin{lemma}[First-return counting]\label{lem:counting}
Let $F:X\to X$ be a map on a finite set $|X|=N$, and let $L\subseteq X$.
Assume that for every $x\in L$ the first return time
\[
\rho(x)=\min\{t\ge 1:F^t(x)\in L\}
\]
exists, and write
\[
T(x)=F^{\rho(x)}(x)\in L
\]
for the induced first-return map.
If
\begin{enumerate}[label=\textup{(\roman*)}]
\item $T$ is a single $|L|$-cycle on $L$, and
\item $\sum_{x\in L}\rho(x)=N$,
\end{enumerate}
then $F$ is a single $N$-cycle on $X$.
\end{lemma}

\begin{remark}
Lemma~\ref{lem:counting} is a discrete analogue of the Poincar\'e first-return map in dynamical systems: it reduces the global cycle structure of~$F$ to the dynamics of the induced map on a cross-section.  We use it as the common final step in all three Route~E color proofs.
\end{remark}

\begin{proof}
Pick $x_0\in L$.
Because $T$ is a cycle on all of $L$, the return points along the $F$-orbit of $x_0$ are
\[
x_0,\ T(x_0),\ T^2(x_0),\ \dots,\ T^{|L|-1}(x_0),\ x_0,
\]
and every point of $L$ appears exactly once before the sequence closes.

Set
\[
N_0:=\sum_{r=0}^{|L|-1}\rho\bigl(T^r(x_0)\bigr).
\]
By construction, after $N_0$ applications of $F$ one returns to $x_0$.
So the $F$-orbit of $x_0$ is a cycle of length $N_0$.
The hypothesis gives
\[
N_0=\sum_{x\in L}\rho(x)=N.
\]
Hence one $F$-cycle already has size $N=|X|$, so it contains all points of $X$.
Thus $F$ is a single $N$-cycle.
\end{proof}

Across the three colors we choose coordinates adapted to the \emph{bulk branch} of the return map---the branch that applies at all but boundedly many points of $P_0$.
In each coordinate system the generic step advances one coordinate (the \emph{clock}) uniformly while preserving the other (the \emph{lane}).
The finitely many non-generic branches are the \emph{stalls} introduced above: at a stall the lane shifts by a bounded amount, and the clock may gain or lose an extra unit.
In the orbit displays below, $\xrightarrow{\bulk}$ or $\xrightarrow{\bulk^t}$ denotes one or $t$ consecutive generic steps; $\xrightarrow{+1}$ or $\xrightarrow{+2}$ denotes a stall that increases the lane by $1$ or $2$.

Since the number of stalls per orbit is bounded independently of $m$, the two-dimensional dynamics on $P_0$ reduce to one-dimensional first-return maps on explicit transversals.
On those transversals, most lanes follow the same arithmetic rule; only $O(1)$ boundary lanes have different itineraries.

\begin{lemma}[bulk coordinates for the three colors]\label{lem:routeE-bulkcoords}
Define linear maps $\Phi_c:P_0\cong \Zm^2\to \Zm^2$ by
\[
\Phi_1(i,k)=(i-k,\ k),
\]
\[
\Phi_2(i,k)=(2i+k,\ -i-k),
\]
\[
\Phi_0(i,k)=(i+2k,\ k).
\]
Let
\[
b_1=(1,1),\qquad b_2=(1,-2),\qquad b_0=(-2,1).
\]
Then each $\Phi_c$ is invertible over $\Zm$, and in the corresponding $(u,t)$-coordinates the bulk move $z\mapsto z+b_c$ becomes
\[
(u,t)\mapsto (u,t+1).
\]
\end{lemma}

\begin{proof}
The matrices of $\Phi_1,\Phi_2,\Phi_0$ are
\[
M_1=\begin{pmatrix}1&-1\\0&1\end{pmatrix},\qquad
M_2=\begin{pmatrix}2&1\\-1&-1\end{pmatrix},\qquad
M_0=\begin{pmatrix}1&2\\0&1\end{pmatrix}.
\]
Their determinants are $1,-1,1$, so each is a unit in $\Zm$ and each map is invertible.
A direct multiplication gives
\[
M_1b_1=(0,1),\qquad M_2b_2=(0,1),\qquad M_0b_0=(0,1),
\]
which is exactly the claim.
\end{proof}

\begin{lemma}[finite-defect normal form for Route~E]\label{lem:routeE-finite-defect}
For each color $c\in\{0,1,2\}$, let
\[
\widehat R_c:=\Phi_c\circ R_c\circ \Phi_c^{-1}.
\]
Then there is a defect set $E_c\subseteq \Zm^2$, contained in a finite union of affine lines together with finitely many isolated points, such that
\[
\widehat R_c(u,t)=(u,t+1)\qquad\text{for }(u,t)\notin E_c.
\]
More precisely:
\begin{enumerate}[label=\textup{(\roman*)},leftmargin=2.5em]
\item For color $1$, every defect step is one of
\[
(u,t)\mapsto (u+1,t),\qquad (u,t)\mapsto (u+2,t),
\]
and the defect set is contained in
\[
u+t=0,\qquad u+2t=m-1,
\]
and, when $m\equiv 4\pmod 6$, also
\[
u+t=1,
\]
together with finitely many isolated points.

\item For color $2$, every defect step is one of
\[
(u,t)\mapsto (u-1,t+2),\quad (u,t)\mapsto (u+1,t+1),\quad (u,t)\mapsto (u-1,t+1),
\]
\[
(u,t)\mapsto (u-2,t+2),\quad (u,t)\mapsto (u+1,t),
\]
and the defect set is contained in
\[
t=1,\qquad t=m-1,\qquad u+2t=0,\qquad u+t=m-1,
\]
together with finitely many isolated points.

\item For color $0$, every defect step is one of
\[
(u,t)\mapsto (u-2,t),\quad (u,t)\mapsto (u+3,t+3),\quad (u,t)\mapsto (u+2,t+2),
\]
\[
(u,t)\mapsto (u-1,t),\quad (u,t)\mapsto (u+1,t+2),\quad (u,t)\mapsto (u+1,t+1),
\]
and the defect set is contained in
\[
t=0,\qquad u=t+1,
\]
and, when $m\equiv 4\pmod 6$, also
\[
u=1+2t,
\]
together with finitely many isolated points.
\end{enumerate}
\end{lemma}

\begin{proof}
We compute directly from the piecewise formulas for $R_0,R_1,R_2$ recorded in Appendix~\ref{app:routeE}.

For color $1$, the generic branch is $(i,k)\mapsto(i+1,k+1)$, while the two exceptional branches are $(i,k)\mapsto(i+1,k)$ and $(i,k)\mapsto(i+2,k)$.
Under $\Phi_1(i,k)=(u,t)=(i-k,k)$ these become
\[
(i+1,k+1)\mapsto (u,t+1),\qquad (i+1,k)\mapsto (u+1,t),\qquad (i+2,k)\mapsto (u+2,t).
\]
The supporting conditions in Appendix~\ref{app:routeE} are $i=0$, $i=1$, and $i+k\equiv m-1$, plus finitely many exceptional boundary points.
These are exactly the affine conditions
\[
i=0 \iff u+t=0,\qquad i=1 \iff u+t=1,\qquad i+k\equiv m-1 \iff u+2t=m-1.
\]
This proves \textup{(i)}.

For color $2$, the generic branch is $(i,k)\mapsto(i+1,k-2)$.
The exceptional branches listed in Appendix~\ref{app:routeE} are
\[
(i+1,k-3),\qquad (i+2,k-3),\qquad (i,k-1),\qquad (i,k-2),\qquad (i+1,k-1).
\]
Under $\Phi_2(i,k)=(u,t)=(2i+k,-i-k)$ one obtains
\[
(i+1,k-2)\mapsto (u,t+1),
\]
\[
(i+1,k-3)\mapsto (u-1,t+2),\qquad (i+2,k-3)\mapsto (u+1,t+1),
\]
\[
(i,k-1)\mapsto (u-1,t+1),\qquad (i,k-2)\mapsto (u-2,t+2),\qquad (i+1,k-1)\mapsto (u+1,t).
\]
The supporting conditions in Appendix~\ref{app:routeE} are $i+k\equiv m-1$, $i+k\equiv 1$, $k=0$, and $i=m-1$, together with finitely many exceptional points.
These become
\[
i+k\equiv m-1 \iff t=1,
\]
\[
i+k\equiv 1 \iff t=m-1,
\]
\[
k=0 \iff u+2t=0,
\]
\[
i=m-1 \iff u+t=m-1.
\]
This proves \textup{(ii)}.

For color $0$, the generic branch is $(i,k)\mapsto(i-2,k+1)$.
The exceptional branches listed in Appendix~\ref{app:routeE} are
\[
(i-2,k),\qquad (i-3,k+3),\qquad (i-2,k+2),\qquad (i-1,k),\qquad (i-3,k+2),\qquad (i-1,k+1).
\]
Under $\Phi_0(i,k)=(u,t)=(i+2k,k)$ these become
\[
(i-2,k+1)\mapsto (u,t+1),
\]
\[
(i-2,k)\mapsto (u-2,t),\qquad (i-3,k+3)\mapsto (u+3,t+3),\qquad (i-2,k+2)\mapsto (u+2,t+2),
\]
\[
(i-1,k)\mapsto (u-1,t),\qquad (i-3,k+2)\mapsto (u+1,t+2),\qquad (i-1,k+1)\mapsto (u+1,t+1).
\]
The supporting conditions in Appendix~\ref{app:routeE} are $k=0$, $i+k\equiv 1$, and, in Case~II, $i=1$, together with finitely many exceptional points.
These become
\[
k=0 \iff t=0,
\]
\[
i+k\equiv 1 \iff u=t+1,
\]
\[
i=1 \iff u=1+2t.
\]
This proves \textup{(iii)}.
\end{proof}

\begin{remark}[origin of the Case~I\,/\,II split]
\label{rem:mod6-split}
The distinction between Case~I ($m\equiv 0,2\pmod 6$) and Case~II ($m\equiv 4\pmod 6$) that recurs in the color-by-color proofs below originates here.
When $m\equiv 0,2\pmod 6$ certain defect lines are absent---for instance, the line $u+t=1$ in part~\textup{(i)} and the line $u=1+2t$ in part~\textup{(iii)} appear only in Case~II---so the orbit encounters fewer defect components and the first-return formulas are simpler.
When $m\equiv 4\pmod{6}$ these extra defect lines are present, which introduces additional stall patterns and accounts for the distinct case analysis.
\end{remark}

\begin{lemma}[defect-itinerary reduction]\label{lem:routeE-itinerary}
Let $F:\Zm^2\to\Zm^2$ be a map of the form
\[
F(z)=
\begin{cases}
G(z):=(u,t+1), & z\notin E,\\
D_\alpha(z):=(u+\Delta_\alpha,\ t+\tau_\alpha), & z\in E_\alpha,
\end{cases}
\]
where $E=\bigsqcup_\alpha E_\alpha$ is a finite disjoint union of defect components and each $D_\alpha$ has fixed increment $(\Delta_\alpha,\tau_\alpha)$.
Let
\[
L:=\{(u,0):u\in\Zm\}.
\]
Assume every point of $L$ has a first return to $L$.
Then for every start $(u,0)\in L$, the orbit segment before first return has a unique factorization
\[
G^{a_0}D_{\alpha_1}G^{a_1}D_{\alpha_2}\cdots D_{\alpha_r}G^{a_r},
\]
where each $a_j\ge 0$ and every point of each generic block lies outside $E$.
If
\[
F^{\rho(u)}(u,0)=(T(u),0),
\]
then
\[
T(u)=u+\sum_{j=1}^r \Delta_{\alpha_j}\pmod m,
\qquad
\rho(u)=\sum_{j=0}^r a_j+r.
\]
\end{lemma}

\begin{proof}
Because $F$ is deterministic, once an orbit point lies outside $E$ it is forced to follow the generic branch $G$ until it either returns to $L$ or first meets one of the defect components.
Starting from $(u,0)$, let $a_0$ be the maximal integer such that
\[
(u,0),\ G(u,0),\ \dots,\ G^{a_0-1}(u,0)
\]
all lie outside $E$.
If $G^{a_0}(u,0)\in L$, then there are no defect steps and the claim is immediate.
Otherwise $G^{a_0}(u,0)$ lies in a unique defect component $E_{\alpha_1}$, so the next step is forced to be $D_{\alpha_1}$.
Repeating the same argument recursively gives the displayed factorization, and uniqueness follows because there is never any branch choice.

The generic branch $G$ leaves the $u$-coordinate unchanged, while each defect step $D_{\alpha_j}$ changes $u$ by exactly $\Delta_{\alpha_j}$.
Therefore the net lane change from $(u,0)$ to $(T(u),0)$ is
\[
\sum_{j=1}^r \Delta_{\alpha_j},
\]
which proves the formula for $T(u)$.
The total number of steps is the number of generic steps plus the number of defect steps, giving
\[
\rho(u)=\sum_{j=0}^r a_j+r.
\]
\end{proof}

Combining Lemma~\ref{lem:routeE-finite-defect} with Lemma~\ref{lem:routeE-itinerary} reduces each Route~E cycle proof to a one-dimensional first-return problem on the transversal $L=\{(u,0):u\in\Zm\}$: one lists the defect components encountered by the bulk ray from $(u,0)$ before first return, reads off the lane increment and return time, and checks that the induced map $T$ is a single $m$-cycle with $\sum_u\rho(u)=m^2$.

\begin{table}[t]
\centering
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabularx}{\textwidth}{@{}l l l >{\raggedright\arraybackslash}X@{}}
\toprule
color & bulk coordinates $(u,t)$ & bulk branch & defect support / transversal \\
\midrule
$1$ & $(i-k,\,k)$ & $(u,t)\mapsto(u,t+1)$ & defects on $u+t=0$, $u+2t=m-1$, and in Case~II also $u+t=1$; transversal $t=0$ \\
$2$ & $(2i+k,\,-i-k)$ & $(u,t)\mapsto(u,t+1)$ & defects on $t=1$, $t=m-1$, $u+2t=0$, $u+t=m-1$; transversal $t=0$ \\
$0$ & $(i+2k,\,k)$ & $(u,t)\mapsto(u,t+1)$ & defects on $t=0$, $u=t+1$, and in Case~II also $u=1+2t$; transversal $t=0$ \\
\bottomrule
\end{tabularx}
\caption{Route~E at a glance: adapted bulk coordinates, generic branch, and affine defect support for the three colors. The main text works out color $2$ explicitly, while the appendix computes the remaining mod-$6$-dependent first-return data on the listed transversals.}
\label{tab:routeE-summary}
\end{table}

\begin{figure}[t]
\centering
\begin{minipage}[t]{0.32\textwidth}
\centering
\begin{tikzpicture}[x=0.62cm,y=0.62cm,>=stealth,line cap=round,line join=round]
\draw[thick] (0,0) rectangle (5,5);
\foreach \x in {1.2,2.5,3.8} {\draw[->] (\x,1.0) -- (\x,2.2);}
\draw[very thick] (0,5) -- (5,0);
\draw[very thick,dashed] (0,3.8) -- (5,1.3);
\draw[very thick,dotted] (0.8,5) -- (5,0.8);
\node[font=\scriptsize,anchor=south west] at (0.05,5) {$u+t=0$};
\node[font=\scriptsize,anchor=south west] at (0.05,3.8) {$u+2t=m-1$};
\node[font=\scriptsize,anchor=south west] at (0.85,4.6) {$u+t=1$};
\end{tikzpicture}

\smallskip
{\footnotesize Color $1$ in $(u,t)$-coordinates}
\end{minipage}\hfill
\begin{minipage}[t]{0.32\textwidth}
\centering
\begin{tikzpicture}[x=0.62cm,y=0.62cm,>=stealth,line cap=round,line join=round]
\draw[thick] (0,0) rectangle (5,5);
\foreach \x/\y in {0.8/4.1,1.8/3.1,2.8/2.1} {\draw[->] (\x,\y) -- ++(0.9,-0.9);}
\draw[very thick] (0,4.1) -- (5,4.1);
\draw[very thick,dashed] (0,0.9) -- (5,0.9);
\draw[very thick,dotted] (0,0) -- (5,5);
\draw[very thick,dash dot] (4.1,0) -- (4.1,5);
\node[font=\scriptsize,anchor=south west] at (0.05,4.15) {$y=m-1$};
\node[font=\scriptsize,anchor=south west] at (0.05,0.95) {$y=1$};
\node[font=\scriptsize,anchor=south west] at (2.55,2.55) {$y=x$};
\node[font=\scriptsize,anchor=south west] at (4.15,4.45) {$x=m-1$};
\end{tikzpicture}

\smallskip
{\footnotesize Color $2$ in $(x,y)=(i,i+k)$}
\end{minipage}\hfill
\begin{minipage}[t]{0.32\textwidth}
\centering
\begin{tikzpicture}[x=0.62cm,y=0.62cm,>=stealth,line cap=round,line join=round]
\draw[thick] (0,0) rectangle (5,5);
\foreach \x in {1.2,2.5,3.8} {\draw[->] (\x,0.9) -- (\x,2.0);}
\draw[very thick] (0,0.9) -- (5,0.9);
\draw[very thick,dashed] (0.9,0) -- (5,4.1);
\draw[very thick,dotted] (1.2,0) -- (5,1.9);
\node[font=\scriptsize,anchor=south west] at (0.05,0.95) {$y=0$};
\node[font=\scriptsize,anchor=south west] at (2.9,2.35) {$x=y+1$};
\node[font=\scriptsize,anchor=south west] at (2.4,1.05) {$x=1+2y$};
\end{tikzpicture}

\smallskip
{\footnotesize Color $0$ in $(x,y)=(i+2k,k)$}
\end{minipage}
\caption{Schematic defect geometry on $P_0$. The left and right panels use the bulk coordinates of Lemma~\ref{lem:routeE-bulkcoords}; the center panel (color~$2$) uses the working frame $(x,y)=(i,\,i+k)$. In each panel the arrows indicate the generic bulk branch, while the solid/dashed/dotted lines represent the affine defect families from Lemma~\ref{lem:routeE-finite-defect}. The isolated boundary-point corrections sit on these lines and act as the splice points described in Remark~\ref{rem:routeE-design}.}
\label{fig:routeE-defects}
\end{figure}

\begin{remark}[$R_2$ is universal]\label{rem:R2-universal}
The return-map formula for color~$2$ is the same for all even $m\ge 6$, independent of $m\bmod 6$; only colors~$0$ and~$1$ require a Case~I\,/\,II split.
For this reason, the main text works out color~$2$ completely before the appendix turns to the genuinely mod-$6$-dependent colors.
\end{remark}

\begin{example}[Color $2$ for $m=6$]\label{ex:m6-color2}
We illustrate the return-map machinery on the smallest admissible even value $m=6$.
In the working frame $(x,y)=(i,\,i+k)$ on $P_0$, the generic (bulk) step is $G\colon(x,y)\mapsto(x+1,y-1)$, and the transversal is $L_2=\{(x,0):x\in\mathbb Z_6\}$.

Starting from lane $x=0$:
\[
(0,0)\xrightarrow{E}(1,0).
\]
This single defect step already returns to $L_2$, giving $T_2(0)=1$ and $\rho_2(0)=1$.

Starting from lane $x=1$, the orbit crosses two defect lines before returning:
\[
(1,0)\xrightarrow{G}(2,5)\xrightarrow{C}(2,4)\xrightarrow{G}(3,3)
\xrightarrow{D}(3,1)\xrightarrow{B}(5,0).
\]
The defect sequence is $C$ (line $y=m{-}1$), then $D$ (diagonal $y=x$), then $B$ (line $y=1$).
Return data: $T_2(1)=5$, $\rho_2(1)=5$.

For the remaining lanes, the same method gives $T_2(2)=0$ with $\rho_2(2)=12=2m$, and $T_2(x)=x-1$ with $\rho_2(x)=6=m$ for $x\in\{3,4,5\}$.
The first-return map $T_2$ is therefore the $6$-cycle
\[
0\to 1\to 5\to 4\to 3\to 2\to 0,
\]
and the return-time sum is $1+5+12+6+6+6=36=6^2=m^2$.
Lemma~\ref{lem:counting} now implies that $R_2$ is a single $36$-cycle on $P_0$.
\end{example}

\subsection{The color-$2$ return map}\label{sec:even-r2}
We now carry out the color-$2$ computation in full.
This is the cleanest representative case because its return map does not depend on $m\bmod 6$.
For color~$2$ it is convenient to work directly in the coordinates $(x,y)=(i,\,i+k)$ rather than the bulk frame of Lemma~\ref{lem:routeE-bulkcoords}; in this frame the generic branch becomes $G\colon(x,y)\mapsto(x+1,y-1)$ and the defect lines are those shown in the center panel of Figure~\ref{fig:routeE-defects}.

\begin{lemma}\label{lem:R2-main-xy}
In the coordinates $(x,y)=(i,i+k)$ on $P_0$, the associated Route~E $m$-step map is
\[
R_2(x,y)=
\begin{cases}
(x+1,y-2),&(x,y)=(2,2),\\
(x+2,y-1),&y=1,\ (x,y)\notin\{(1,1),(m-1,1)\},\\
(x,y-1),&\bigl(y=m-1,\ x\neq m-1\bigr)\ \text{or}\ (x,y)=(m-1,m-2),\\
(x,y-2),&\bigl(y=x,\ x\notin\{0,2,m-1\}\bigr)\ \text{or}\ (x,y)=(m-1,0),\\
(x+1,y),&\bigl(x=m-1,\ y\notin\{0,m-2\}\bigr)\ \text{or}\ (x,y)=(0,0),\\
(x+1,y-1),&\text{otherwise}.
\end{cases}
\]
All arithmetic is modulo $m$.
\end{lemma}

\begin{proof}
Substitute $i=x$ and $k=y-x$ into the closed-form formula for $R_2$ from Appendix~\ref{app:routeE} and simplify each branch.
The six displayed moves are exactly the transformed images of the six original branches.
\end{proof}

For the orbit traces below write
\[
G(x,y)=(x+1,y-1)
\]
for the generic branch of Lemma~\ref{lem:R2-main-xy}.
The other five displayed branches will be referred to as $A,B,C,D,E$ in the order listed in the lemma:
$A$ applies at the isolated point $(2,2)$;
$B$ is the defect on the line $y=1$;
$C$ is the defect on $y=m-1$;
$D$ is the defect on $y=x$;
$E$ is the defect on $x=m-1$.

The transversal is
\[
L_2=\{(x,0):x\in \Zm\}.
\]

\begin{proposition}\label{prop:R2-main-data}
For every even $m\ge 6$, the first-return map on $L_2$ is
\[
T_2(x)=
\begin{cases}
1,&x=0,\\
m-1,&x=1,\\
0,&x=2,\\
x-1,&3\le x\le m-1,
\end{cases}
\]
with return times
\[
\rho_2(x)=
\begin{cases}
1,&x=0,\\
m-1,&x=1,\\
2m,&x=2,\\
m,&3\le x\le m-1.
\end{cases}
\]
\end{proposition}

\begin{proof}
We treat the boundary values $x\in\{0,1,2,4,m-2,m-1\}$ individually and then give the generic odd-$x$ and even-$x$ formulas.
Fix $x\in \Zm$ and start from $(x,0)\in L_2$.

\smallskip
\noindent\textbf{Case $x=0$.}
The point $(0,0)$ is branch $E$, so
\[
(0,0)\xrightarrow{E}(1,0).
\]
Hence $T_2(0)=1$ and $\rho_2(0)=1$.

\smallskip
\noindent\textbf{Case $x=1$.}
The orbit is
\[
(1,0)\xrightarrow{G}(2,m-1)\xrightarrow{C}(2,m-2)
\xrightarrow{G^{\frac m2-2}}\left(\frac m2,\frac m2\right)
\xrightarrow{D}\left(\frac m2,\frac m2-2\right)
\xrightarrow{G^{\frac m2-3}}(m-3,1)\xrightarrow{B}(m-1,0).
\]
Between the displayed defect points, no other special branch can occur:
before reaching $\left(\frac m2,\frac m2\right)$ the orbit has $x<m-1$, $y\notin\{0,1,m-1\}$, and meets the diagonal only at the displayed point; after the $D$-move the next special point is $(m-3,1)$.
Therefore
\[
T_2(1)=m-1,\qquad
\rho_2(1)=1+1+\left(\frac m2-2\right)+1+\left(\frac m2-3\right)+1=m-1.
\]

\smallskip
\noindent\textbf{Case $x=2$.}
The orbit is
\[
(2,0)\xrightarrow{G}(3,m-1)\xrightarrow{C}(3,m-2)
\xrightarrow{G^{m-4}}(m-1,2)\xrightarrow{E}(0,2)\xrightarrow{G}(1,1)
\xrightarrow{D}(1,m-1)\xrightarrow{C}(1,m-2)
\xrightarrow{G^{m-3}}(m-2,1)\xrightarrow{B}(0,0).
\]
Again, every point between the displayed defects is generic:
before $(m-1,2)$ the orbit avoids the diagonal and the line $y=1$;
after $(0,2)$ the only diagonal hit is $(1,1)$;
after $(1,m-2)$ the next defect is $(m-2,1)$.
Hence
\[
T_2(2)=0,\qquad
\rho_2(2)=1+1+(m-4)+1+1+1+1+(m-3)+1=2m.
\]

\smallskip
\noindent\textbf{Case $x$ odd, $3\le x\le m-3$.}
Set
\[
d=\frac{x+m-1}{2}.
\]
Then the orbit is
\[
(x,0)\xrightarrow{G}(x+1,m-1)\xrightarrow{C}(x+1,m-2)
\xrightarrow{G^{\frac{m-x-3}{2}}}(d,d)\xrightarrow{D}(d,d-2)
\xrightarrow{G^{\frac{m-x-1}{2}}}(m-1,x-2)\xrightarrow{E}(0,x-2)
\xrightarrow{G^{x-3}}(x-3,1)\xrightarrow{B}(x-1,0).
\]
The first generic block runs on the line of ordinary sum $x+m-1$, which meets the diagonal exactly at $(d,d)$.
After the $D$-move the orbit runs on the line of ordinary sum $x+m-3$, and its next defect is $(m-1,x-2)$.
After the $E$-move the orbit runs on the line of ordinary sum $x-2$, whose first defect is $(x-3,1)$.
Thus
\[
T_2(x)=x-1,
\]
and
\[
\rho_2(x)=
1+1+\frac{m-x-3}{2}+1+\frac{m-x-1}{2}+1+(x-3)+1
=
m.
\]

\smallskip
\noindent\textbf{Case $x=4$.}
Here the orbit is
\[
(4,0)\xrightarrow{G}(5,m-1)\xrightarrow{C}(5,m-2)
\xrightarrow{G^{m-6}}(m-1,4)\xrightarrow{E}(0,4)
\xrightarrow{G^2}(2,2)\xrightarrow{A}(3,0).
\]
The displayed formula also covers the border case $m=6$, where the block $G^{m-6}$ is empty.
Hence
\[
T_2(4)=3,\qquad \rho_2(4)=1+1+(m-6)+1+2+1=m.
\]

\smallskip
\noindent\textbf{Case $x$ even, $6\le x\le m-4$.}
Set
\[
e=\frac{x}{2}.
\]
Then the orbit is
\[
(x,0)\xrightarrow{G}(x+1,m-1)\xrightarrow{C}(x+1,m-2)
\xrightarrow{G^{m-x-2}}(m-1,x)\xrightarrow{E}(0,x)
\xrightarrow{G^e}(e,e)\xrightarrow{D}(e,e-2)
\xrightarrow{G^{e-3}}(x-3,1)\xrightarrow{B}(x-1,0).
\]
Before $(m-1,x)$ there is no diagonal hit because the first generic block lies on a line of odd ordinary sum.
After the $E$-move the orbit lies on the line of ordinary sum $x$, whose first diagonal point is $(e,e)$.
After the $D$-move the orbit lies on the line of ordinary sum $x-2$, whose next defect is $(x-3,1)$.
Therefore
\[
T_2(x)=x-1,
\]
and
\[
\rho_2(x)=
1+1+(m-x-2)+1+e+1+(e-3)+1
=
m.
\]

\smallskip
\noindent\textbf{Case $x=m-2$ (so necessarily $m\ge 8$).}
The orbit is
\[
(m-2,0)\xrightarrow{G}(m-1,m-1)\xrightarrow{E}(0,m-1)\xrightarrow{C}(0,m-2)
\xrightarrow{G^{\frac{m-2}{2}}}\left(\frac{m-2}{2},\frac{m-2}{2}\right)
\xrightarrow{D}\left(\frac{m-2}{2},\frac{m-6}{2}\right)
\xrightarrow{G^{\frac{m-8}{2}}}(m-5,1)\xrightarrow{B}(m-3,0).
\]
Hence
\[
T_2(m-2)=m-3,\qquad
\rho_2(m-2)=1+1+1+\frac{m-2}{2}+1+\frac{m-8}{2}+1=m.
\]

\smallskip
\noindent\textbf{Case $x=m-1$.}
The orbit is
\[
(m-1,0)\xrightarrow{D}(m-1,m-2)\xrightarrow{C}(m-1,m-3)\xrightarrow{E}(0,m-3)
\xrightarrow{G^{m-4}}(m-4,1)\xrightarrow{B}(m-2,0).
\]
So
\[
T_2(m-1)=m-2,\qquad
\rho_2(m-1)=1+1+1+(m-4)+1=m.
\]

The seven cases exhaust $x\in \Zm$, and the displayed formulas follow.
\end{proof}

\begin{corollary}[color $2$ is Hamilton for Route~E]\label{cor:R2-main}
For every even $m\ge 6$, the associated map $R_2$ is a single $m^2$-cycle on $P_0$.
\end{corollary}

\begin{proof}
By Proposition~\ref{prop:R2-main-data}, the induced lane map follows the cyclic order
\[
(0,1\mid m-1,m-2,\dots,2),
\]
namely $T_2(0)=1$, $T_2(1)=m-1$, $T_2(2)=0$, and $T_2(x)=x-1$ for $3\le x\le m-1$.
Hence $T_2$ is a single $m$-cycle.
Also
\[
\sum_{x\in \Zm}\rho_2(x)=1+(m-1)+2m+(m-3)m=m^2.
\]
Lemma~\ref{lem:counting} now implies that $R_2$ is a single $m^2$-cycle.
\end{proof}

\begin{theorem}[Route~E for even $m\ge 6$]\label{thm:routeE-even-main}
For every even integer $m\ge 6$, the Route~E direction assignment (Definition~\ref{def:routeE}) is a valid coloring of $D(m)$, and its three color classes are directed Hamilton cycles.
\end{theorem}

\begin{proof}
Proposition~\ref{prop:routeE-transducer} reduces the derivation of the return maps to a finite low-layer table, and Appendix~\ref{app:routeE-derivation} carries out that evaluation, yielding the closed-form formulas for $R_0,R_1,R_2$ recorded in Appendix~\ref{app:routeE}.
Lemmas~\ref{lem:routeE-bulkcoords}, \ref{lem:routeE-finite-defect}, and \ref{lem:routeE-itinerary} show that each return map is a finite-defect skew product whose cycle structure is governed by a one-dimensional first-return map on a transversal.
Corollary~\ref{cor:R2-main} proves in the main text that $R_2$ is a single $m^2$-cycle on $P_0$.
The color-$2$ argument serves as the template for the remaining two colors: in each case, one passes to adapted bulk coordinates (Lemma~\ref{lem:routeE-bulkcoords}), identifies the affine defect lines (Lemma~\ref{lem:routeE-finite-defect}), and reads off the induced lane map from the defect itinerary (Lemma~\ref{lem:routeE-itinerary}).
Colors $0$ and $1$ differ from color $2$ only in their bulk coordinate systems, their defect geometry, and the resulting mod-$6$ case split.
Appendix~\ref{app:routeE-cycle} carries out this analysis: Corollary~\ref{cor:R1-all} concludes that $R_1$ is a single $m^2$-cycle, and Corollary~\ref{cor:R0-all} does the same for $R_0$.
Lemma~\ref{lem:counting} therefore implies that each $R_c$ is a single $m^2$-cycle on $P_0$.
In particular, each $R_c$ is bijective, so Lemma~\ref{lem:return-validity} implies that the Route~E direction assignment is a valid coloring.
Then Lemma~\ref{lem:return} shows that each color map is a directed Hamilton cycle on $V$.
Because every Route~E direction triple is a permutation of $(0,1,2)$, the three Hamilton cycles are arc-disjoint and together cover all arcs of $D(m)$.
\end{proof}


\begin{remark}[Finite-defect design principle]
The Route~E construction shares the same return-map framework as the odd affine construction---Hamiltonicity of the full color map on $V$ is reduced to the cycle structure of a return map on $P_0$, which is further reduced to a one-dimensional first-return problem---but the execution is considerably more intricate.
Where the odd case enjoys a single uniform return-map formula, the even return maps have piecewise definitions with an $m\bmod 6$ case split and boundary-point exceptions.
Its low-layer modifications are supported on a bounded number of affine lines and exceptional points inside the planes $S=0,1,2$.
Away from this defect set, the return maps behave like simple bulk translations on $P_0$.
\end{remark}
\section{The finite case $m=4$}\label{sec:m4}

Route~E is designed for $m\ge 6$; the case $m=4$ is handled by a separate finite witness.
Appendix~\ref{app:m4} prints the complete direction table, from which the three Hamilton cycles can be read off by finite iteration; a machine-readable supplement records the orbits explicitly.

\begin{proposition}[The finite case $m=4$]\label{prop:m4-main}
The graph $D(4)$ admits a decomposition into three directed Hamilton cycles.
\end{proposition}

\begin{proof}
Appendix~\ref{app:m4} records a complete direction table for a coloring of $D(4)$.
Starting from that table, each color orbit is obtained by finite iteration from any chosen initial vertex; in the accompanying machine-readable supplement \texttt{m4\_witness.json} we record the three orbits starting from $(0,0,0)$, and the script \texttt{verify\_m4\_witness.py} checks that each one has length $64$ and agrees with the displayed table.
Thus the finite verification splits into two transparent steps: the table shows that every vertex sees a permutation of the three allowed directions, hence defines a valid direction assignment, and the extracted orbit data show that each color map is a single $64$-cycle on $V$, hence a bijection; together these establish a valid coloring.
Therefore the three color classes are Hamilton cycles and form a Hamilton decomposition of $D(4)$.
\end{proof}


\section{Completion for all $m\ge 3$}\label{sec:completion}

\begin{proof}[Proof of Theorem~\ref{thm:main-all}]
If $m$ is odd, apply Theorem~\ref{thm:odd-hamilton}.
If $m=4$, apply Proposition~\ref{prop:m4-main}.
If $m$ is even and $m\ge 6$, apply Theorem~\ref{thm:routeE-even-main}.
These cases exhaust all integers $m\ge 3$.
\end{proof}

\section{Discussion}\label{sec:discussion}

The odd and even constructions are fundamentally different.
For odd $m$, five Kempe swaps from the canonical coloring yield affine skew-product return maps whose single-cycle property follows from elementary modular arithmetic.
For even $m$, the sign-product invariant rules out any Kempe-from-canonical approach, and the proof instead relies on the explicit Route~E direction assignment, whose modifications are concentrated on the three lowest layers.
The infinite-family proofs are purely combinatorial; the boundary case $m=4$ is certified by an explicit finite witness (Appendix~\ref{app:m4}).

Several natural questions remain.
First, it would be desirable to unify the Route~E case analysis---currently split by the residue of $m$ modulo~$6$---into a single conceptual argument.
At the level of the induced lane maps, the present proof shows that the low-layer defects have a single role: they splice a bounded number of bulk arithmetic lane families into one cyclic order.
A shorter derivation of those splice rules, or a structural explanation of why the mod-$6$ split is unavoidable, would both be welcome.
Second, the parity-obstruction viewpoint motivates the search for analogous invariants and obstruction-breaking constructions in higher-dimensional directed tori $\vec C_m^{\,d}$ and in broader families of Cayley digraphs.
In particular, does Corollary~\ref{cor:parity-barrier} generalize to higher dimensions, and if so, does the obstruction admit a similar low-layer fix?

\section*{Acknowledgments}
The author has no acknowledgments to report.

\section*{Conflict of interest}
The author declares no conflict of interest.

\section*{Data availability statement}
Data sharing not applicable to this article as no datasets were generated or analysed during the current study.
Supporting verification code and the machine-readable $m=4$ witness referenced in Appendix~\ref{app:verification} are supplied with the submission as supporting information.


\begin{thebibliography}{99}

\bibitem{AlspachBermondSotteau1990}
B.~Alspach, J.-C.~Bermond, and D.~Sotteau,
\newblock Decomposition into cycles I: Hamilton decompositions,
\newblock in {\em Cycles and Rays} (G.~Hahn, G.~Sabidussi, R.~E.~Woodrow, eds.),
NATO ASI Series, vol.~301, Springer, 1990, pp.~9--18.

\bibitem{Bermond1978}
J.-C.~Bermond,
\newblock Hamilton decomposition of graphs, directed graphs and hypergraphs,
\newblock in {\em Advances in Graph Theory}, Annals of Discrete Mathematics 3 (1978), 21--28.

\bibitem{Bogdanowicz2017}
Z.~R.~Bogdanowicz,
\newblock On decomposition of the Cartesian product of directed cycles into cycles of equal lengths,
\newblock {\em Discrete Applied Mathematics} 229 (2017), 148--150.

\bibitem{Bogdanowicz2020}
Z.~R.~Bogdanowicz,
\newblock Identifying Hamilton cycles in the Cartesian product of directed cycles,
\newblock {\em AKCE International Journal of Graphs and Combinatorics} 17 (2020), no.~1, 534--538.

\bibitem{CurranGallian1996}
S.~J.~Curran and J.~A.~Gallian,
\newblock Hamiltonian cycles and paths in Cayley graphs and digraphs --- a survey,
\newblock {\em Discrete Mathematics} 156 (1996), 1--18.

\bibitem{CurranWitte1985}
S.~J.~Curran and D.~Witte,
\newblock Hamilton paths in Cartesian products of directed cycles,
\newblock in {\em Cycles in Graphs}, Annals of Discrete Mathematics 27 (1985), 35--74.

\bibitem{DarijaniMiraftabMorris2022}
I.~Darijani, B.~Miraftab, and D.~Witte Morris,
\newblock Arc-disjoint Hamiltonian paths in Cartesian products of directed cycles,
\newblock {\em Ars Mathematica Contemporanea} 25 (2025), no.~2, P2.10.

\bibitem{Foregger1978}
M.~F.~Foregger,
\newblock Hamiltonian decompositions of products of cycles,
\newblock {\em Discrete Mathematics} 24 (1978), 251--260.

\bibitem{Kotzig1973}
A.~Kotzig,
\newblock Every Cartesian product of two circuits is decomposable into two Hamiltonian circuits,
\newblock Centre de Recherches Math\'ematiques, Montr\'eal, Rapport 233 (1973).

\bibitem{Mohar2006}
B.~Mohar,
\newblock Kempe equivalence of colorings,
\newblock in {\em Graph Theory in Paris}, Trends in Mathematics, Birkh\"auser, 2006, pp.~287--297.

\bibitem{StiebitzEtAl2012}
M.~Stiebitz, D.~Scheide, B.~Toft, and L.~M.~Favrholdt,
\newblock {\em Graph Edge Coloring: Vizing's Theorem and Goldberg's Conjecture},
\newblock Wiley Series in Discrete Mathematics and Optimization, Wiley, 2012.

\bibitem{Stong1991}
R.~Stong,
\newblock Hamilton decompositions of Cartesian products of graphs,
\newblock {\em Discrete Mathematics} 90 (1991), 169--190.

\bibitem{TrotterErdos1978}
W.~T.~Trotter and P.~Erd\H{o}s,
\newblock When the Cartesian product of directed cycles is Hamiltonian,
\newblock {\em Journal of Graph Theory} 2 (1978), 137--142.

\bibitem{WitteGallian1984}
D.~Witte and J.~A.~Gallian,
\newblock A survey: Hamiltonian cycles in Cayley graphs,
\newblock {\em Discrete Mathematics} 51 (1984), 293--304.

\end{thebibliography}

\appendix
\input{d3torus_complete_m_ge_3_editorial_revision_appendices}

\section*{Supporting information}
Additional supporting information can be found in the supporting-information files submitted with this article.
The submission bundle contains the directory \texttt{anc/}, including the machine-readable $m=4$ witness, the Route~E verification scripts, and the supporting-information manifest \texttt{SuppInfo\_manifest.md}.

\end{document}
