\documentclass[11pt]{article}
\usepackage{amsmath,amssymb}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xurl}

\hypersetup{colorlinks=true,linkcolor=blue,urlcolor=blue}
\urlstyle{same}

\title{D5 proof bundle overview\\refined March 2026 snapshot}
\author{}
\date{}

\begin{document}
\maketitle

\section*{Purpose}

This note organizes the refined proof bundle for the present
\(T0\)--\(T4,G1,G2\) slice.
In the current state of the bundle, the front-end theorem block and the
selector-compatibility/transport block are closed, while the post-entry
odometer packet and the final globalization packet are still treated as imported
stable inputs.

\section{What the bundle closes and what it still imports}

\begin{enumerate}[leftmargin=2.6em]
\item \textbf{Closed inside the bundle.}
      The actual-needed tiny-repair core \(T0\), the front-end row theorems
      \(T1,T2,T3\), the canonical entry theorem \(T4\), the graph-side
      selector-compatibility theorem \(G1\), and the cyclic transport theorem
      \(G2\).
\item \textbf{Still imported rather than reproved.}
      The post-entry odometer packet and the final graph-side/globalization
      package used after the entry state has been identified.
\end{enumerate}

\section{Primary entry points}

\begin{enumerate}[leftmargin=2.6em]
\item Refined working manuscript:
      \path{manuscript/d5_251_full_d5_working_manuscript_refined.tex}
\item Previous bundle-ready manuscript snapshot:
      \path{manuscript/d5_250_full_d5_working_manuscript_bundle_ready.tex}
\item Review note for the present cleanup pass:
      \path{overview/d5_251_bundle_review_and_polish_note.md}
\item Compact status summaries:
      \path{overview/d5_246_master_status_summary.md} and
      \path{overview/d5_249_master_status_after_G1.md}
\item Surfaced source bundle used in the surface proofs and reruns:
      \path{external/roundy_d5_endpoint_return_model_bundle_20260321_update_197.tar}
\end{enumerate}

\section{Theorem-to-file map}

\begin{description}[leftmargin=2.2cm,style=nextline]
\item[T0 actual-needed core]
      Main note:
      \path{front_end/notes/d5_230_T0_actual_needed_all_odd_surface_proof_note.tex}.
      Main checker:
      \path{front_end/compute/d5_230_T0_needed_core_checker.py}.
\item[T1 strip closure]
      Main notes:
      \path{front_end/notes/d5_220_sigma23_source_surface_all_m_gt2_closure_note.tex}
      and
      \path{front_end/notes/d5_223_sigma32_surface_closure_note.tex}.
\item[T2 obstruction]
      Main note:
      \path{front_end/notes/d5_228_T2_centered_core_obstruction_note.tex}.
      Small-case addendum:
      \path{front_end/notes/d5_229_T2_T3_m3_small_case_note.md}.
\item[T3 obstruction]
      Main note:
      \path{front_end/notes/d5_225_T3_four_collision_families_note.tex}.
      Small-case addendum:
      \path{front_end/notes/d5_229_T2_T3_m3_small_case_note.md}.
\item[T4 canonical entry]
      Main note:
      \path{front_end/notes/d5_233_T4_canonical_entry_surface_proof_note.tex}.
\item[G2 cyclic transport]
      Main note:
      \path{graph_side/notes/d5_234_graph_side_split_G2_closed_G1_remaining_note.tex}.
\item[G1 explicit splice]
      Main note:
      \path{graph_side/notes/d5_247_G1_explicit_two_swap_splice_note.tex}.
      Load-bearing precursors:
      \path{graph_side/notes/d5_236_G1_finite_defect_splice_principle_note.tex}
      and
      \path{graph_side/notes/d5_245_G1_valid_layer_baseline_and_three_class_splice_note.tex}.
\end{description}

\section{Bundle layout}

\begin{verbatim}
d5_251_complete_proof_bundle/
  README.md
  MANIFEST.csv
  manuscript/
  front_end/
    notes/
    compute/
  graph_side/
    notes/
    compute/
  overview/
  external/
\end{verbatim}

\section{Suggested reading order}

\begin{enumerate}[leftmargin=2.6em]
\item Read the refined manuscript first.
\item Read the review note to see what changed in the present polishing pass.
\item For front-end details, read T0, then the two T1 strip-closure notes,
      then T2/T3, then T4.
\item For graph-side details, read the finite-defect splice principle,
      then the valid-baseline/three-class reduction, then the explicit G1 splice,
      and finally the G2 transport note.
\item Use the JSON/CSV files as exact certificates and replay checkpoints.
\end{enumerate}

\section{Compile note}

The refined manuscript and this overview were both compiled in draft mode with
multiple passes during the present cleanup pass so that cross-references are
resolved in the saved compile report.

\end{document}
