\documentclass[11pt]{article}

\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{microtype}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue
}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\Z}{\mathbb Z}
\newcommand{\Rreg}{R_{\mathrm{reg}}}

\title{A compact reproof of the 081 regular closure theorem}
\author{}
\date{March 2026}

\begin{document}
\maketitle

\begin{abstract}
This note replaces the standalone 081 support note by isolating the exact
regular-closure statement used later in 092 and proving it in a shorter
manuscript form.

The proof has two ingredients:
\begin{enumerate}[label=\textup{(\arabic*)},leftmargin=2.5em]
\item the already frozen regular-window formulas from 095, which control the
      source windows and their endpoint labels;
\item the promoted endpoint witness table
      \texttt{d5\_081\_regular\_union\_endpoint\_table.csv} in \texttt{checks/},
      used in the exact sense advertised by the original 081 note: as an
      actual-lift end-gluing certificate on the true larger regular accessible
      union.
\end{enumerate}

The interval arithmetic is uniform in odd $m$. The raw endpoint certificate in
the bundle is checked on the same range as the old 081 support file, namely
$m=13,15,17,19,21$. This note keeps exactly that support range. It does
\emph{not} re-prove the 077 tail-length reduction.
\end{abstract}

\section{Scope and imported endpoint certificate}

Fix an odd modulus $m$ in the accepted D5 regime and the best-seed channel
\[
R1 \to H_{L1}.
\]

Let
\[
\Rreg = \{1,2,4,5,\dots,m-1\}
\]
be the regular source labels.

We use the regular window formulas already frozen in 095:
\begin{itemize}[leftmargin=2em]
\item source-$u$ start label
      \[
      s_u = m((u+2) \bmod m) - 1;
      \]
\item source-$u$ terminal label
      \[
      e_u = m(u-1)-2;
      \]
\item source-$u$ window
      \[
      I_u = [s_u,e_u],
      \]
      a contiguous interval of length $m(m-3)$ in $\Z/m^2\Z$;
\item endpoint shift
      \[
      e_u + 1 = s_{u-3};
      \]
\item grouped successor is exact inside each source window.
\end{itemize}

We also use the promoted regular endpoint witness table
\[
\texttt{checks/d5\_081\_regular\_union\_endpoint\_table.csv}.
\]

In the present note that table is used in the same theorem-level sense stated
by the old 081 support note: each row is an actual-lift endpoint certificate on
the true larger regular accessible union, not merely a quotient-holder summary.

\section{Endpoint arithmetic on regular source windows}

\begin{definition}[regular holders]
For a regular source label $v \in \Rreg$, say that a boundary label $\delta$
has a \emph{source-$v$ regular holder} if $\delta \in I_v$.

A source-$v$ holder is \emph{terminal} when $\delta = e_v$; otherwise it is
\emph{interior}.
\end{definition}

\begin{proposition}[which regular windows contain $e_u$]
Let $u,v \in \Rreg$. Then:
\begin{enumerate}[label=\textup{(\arabic*)},leftmargin=2.5em]
\item $e_u \in I_v$ if and only if $v$ is not one of the three predecessors
      $u-3,u-2,u-1$ (modulo $m$, intersected with $\Rreg$);
\item among the regular holders of $e_u$, the only terminal one is the
      source-$u$ holder.
\end{enumerate}
\end{proposition}

\begin{proof}
Compute in $\Z/m^2\Z$:
\[
e_u - s_v = (m(u-1)-2) - (m((v+2) \bmod m)-1)
          = m((u-v-3) \bmod m) - 1.
\]

Write
\[
t = (u-v-3) \bmod m \in \{0,1,\dots,m-1\}.
\]

Because the source-$v$ window has length $m(m-3)$, the label $e_u$ belongs to
$I_v$ exactly when the difference above is one of the values
\[
m-1,\ 2m-1,\ \dots,\ m(m-3)-1.
\]
Equivalently, $t$ must satisfy
\[
1 \le t \le m-3.
\]
The excluded cases are therefore $t=0,m-2,m-1$, which are exactly the three
predecessors $v=u-3,u-2,u-1$ modulo $m$.

Now $e_u$ is terminal in source $v$ exactly when the difference is the last
point of the interval, namely
\[
e_u - s_v = m(m-3)-1.
\]
That means $t=m-3$, equivalently $v=u$. So the only terminal regular holder of
$e_u$ is the source-$u$ holder.
\end{proof}

\begin{corollary}[containing and continuing regular holders at $e_u$]
For each regular endpoint label $e_u$:
\begin{itemize}[leftmargin=2em]
\item the excluded regular sources are
      \[
      \Rreg \cap \{u-3,u-2,u-1\};
      \]
\item the containing regular holders are
      \[
      \Rreg \setminus (\Rreg \cap \{u-3,u-2,u-1\});
      \]
\item the continuing regular holders are exactly the containing regular holders
      other than $u$.
\end{itemize}
So every containing regular holder of $e_u$ except the source-$u$ holder is
interior and therefore continues to $e_u+1$.
\end{corollary}

\begin{proof}
This is immediate from Proposition~2.2.
\end{proof}

\begin{proposition}[preferred witness pattern in the promoted endpoint table]
A direct row-by-row read of the promoted endpoint table shows the uniform
preferred-witness pattern
\begin{align*}
p(1) &= 2,\\
p(2) &= 4,\\
p(u) &= u+1 \qquad \text{for } 4 \le u \le m-2,\\
p(m-1) &= 1.
\end{align*}
For each row, the recorded successor target is
\[
e_u + 1 = s_{u-3}.
\]
This target is a regular start unless $u=6$, in which case it is the
exceptional start.
\end{proposition}

\begin{proof}
The successor formula is already $e_u+1=s_{u-3}$ from the regular-window
algebra imported from 095. The promoted CSV records exactly the preferred
witness pattern above on every checked row $m=13,15,17,19,21$, and each such
$p(u)$ belongs to the continuing-holder set described in Corollary~2.3. So the
recorded preferred witnesses are always genuine continuing witnesses at the
same endpoint label.
\end{proof}

\section{The reproof replacing 081}

\begin{theorem}[regular endpoint continuation]
For each checked regular endpoint row $e_u$, every actual regular lift at label
$e_u$ glues on the true larger regular accessible union to the continuing class
certified by the preferred witness $p(u)$ from Proposition~2.4.

Consequently, every regular endpoint lift continues to the successor label
\[
e_u + 1 = s_{u-3},
\]
with $u=6$ giving the exceptional start and all other $u$ giving a regular
start.
\end{theorem}

\begin{proof}
By the theorem-level meaning assigned to the promoted endpoint table in the old
081 package, each row of that table is already an actual-lift endpoint
certificate on the true larger regular accessible union. Proposition~2.4
identifies, on each checked row, a preferred same-label continuing witness and
its exact successor target. Therefore every actual regular endpoint lift at
$e_u$ is certified to glue to a continuing witness and hence has a forward
successor to $e_u+1$.
\end{proof}

\begin{theorem}[regular closure theorem; replacement for 081]
On the true larger regular accessible union, every regular realization of every
realized $\delta$ continues.

This is proved on the same checked raw range carried by the promoted endpoint
table:
\[
m = 13,15,17,19,21.
\]
\end{theorem}

\begin{proof}
Take any regular realization and choose one regular holder for its label.

If the realization is interior in that holder, grouped successor exactness
inside the source window gives a forward successor immediately.

If the realization sits at the terminal label of that holder, then its label is
some $e_u$, and Theorem~3.1 gives a forward successor on the true larger
regular accessible union.

So every regular realization continues.
\end{proof}

\begin{corollary}[no regular mixed-status label]
No realized regular $\delta$ has both continuing and terminal regular lifts.
\end{corollary}

\begin{proof}
By Theorem~3.2, every regular lift continues.
\end{proof}

\begin{remark}[what this note does and does not replace]
Theorem~3.2 is exactly the 081 content used in 092:
\begin{itemize}[leftmargin=2em]
\item regular realizations never create a new endpoint sheet;
\item the full odd-$m$ package can no longer fail for a regular reason.
\end{itemize}
This note does \emph{not} separately re-prove the stronger wording
``no regular tail-length ambiguity.'' That belongs to the later 077
fixed-$\delta$ reduction.
\end{remark}

\section{Two representative checked rows}

The promoted CSV contains 75 endpoint rows across the checked moduli
$13,15,17,19,21$. Two representative rows already show the two structural
shapes that matter.

\begin{remark}[$m=13$, ordinary regular handoff]
For $u=4$ one has
\begin{itemize}[leftmargin=2em]
\item $e_4 = 37$;
\item containing regular holders
      \[
      \{4,5,6,7,8,9,10,11,12\};
      \]
\item continuing regular holders
      \[
      \{5,6,7,8,9,10,11,12\};
      \]
\item preferred witness $p(4)=5$;
\item next label $38=s_1$, a regular start.
\end{itemize}
So the unique terminal source-4 holder at 37 is glued to a same-label
continuing witness and therefore does not launch a new regular endpoint sheet.
\end{remark}

\begin{remark}[$m=13$, regular-to-exceptional handoff]
For $u=6$ one has
\begin{itemize}[leftmargin=2em]
\item $e_6 = 63$;
\item containing regular holders
      \[
      \{1,2,6,7,8,9,10,11,12\};
      \]
\item continuing regular holders
      \[
      \{1,2,7,8,9,10,11,12\};
      \]
\item preferred witness $p(6)=7$;
\item next label $64=s_3$, the exceptional start.
\end{itemize}
So even the regular endpoint whose successor lands in the exceptional source
still has a certified continuing witness on the true larger accessible union.
\end{remark}

\section{Package consequence after 096}

After inserting Theorem~3.2 into the cleaned 092 suite, Proposition~4.4 no
longer needs to remain an accepted-support theorem.

Combined with the earlier 094 and 095 reproof notes, the remaining serious
imported support in the cleaned odd-$m$ package is now:
\begin{enumerate}[label=\textup{(\arabic*)},leftmargin=2.5em]
\item 077, the fixed-$\delta$ tail-length reduction;
\item the localized structural 033/062 input already sitting inside the
      first-exit theorem.
\end{enumerate}

So 096 removes 081 from the manuscript-order accepted-support list while
keeping the same honest raw support range as the promoted endpoint table.

\bigskip
\noindent
\textbf{Bottom line.}
The point of 096 is not to reopen the D5 regular sector. It is to state the
regular closure theorem in the exact form later used by 092, and to make the
proof object explicit:
\begin{itemize}[leftmargin=2em]
\item uniform regular-window arithmetic isolates the endpoint rows;
\item the promoted endpoint table certifies actual endpoint gluing there;
\item therefore every regular realization continues;
\item so regular rows can no longer support a new endpoint sheet.
\end{itemize}

\end{document}
